PAWpySeed: A parallelized Python/C package for analyzing DFT wavefunctions in the PAW formalism

### 2018-04-26 docs/paper.tex
update paper for end of Spring 2018 semester

\documentclass[12pt]{article}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{float}
\usepackage{braket}
\usepackage[margin=1.0in]{geometry}
\usepackage[style=chem-acs,backend=bibtex,articletitle=false]{biblatex}
\graphicspath{ {images/} }
\bibliography{mybib}

@@

\title{PAWpySeed: A parallelized Python/C package for analyzing
DFT wavefunctions in the PAW formalism}
\author{Kyle Bystrom\\
        kylebystrom@berkeley.edu\\
		UC Berkeley Department of Materials Science and Engineering}
\date{\today}
\maketitle
\abstract{A parallelized Python and C package has been developed
to perform analysis of density functional theory (DFT) wavefunctions in the
projector augmented wave (PAW) formalism. The primary function of this code
is to calculate the overlap operator of two single-particle Kohn-Sham states
from different periodic structures with the same lattice, which is made nontrivial
by the dependency of PAW augmentation regions on the basis of the structure.
The code is potentially useful for the analysis of point defect levels in
solids because it can determine the composition of defect levels in the basis
of the single-particle states of the bulk system.}

@@

%\subsection{Projector Augmented Wave (PAW) Method}

@@

\subsubsection{Projection onto Basis from Different Structure}
In the pure physical picture of a quantum mechanical system, the eigenfunctions of the time
independent Hamiltonian form a complete basis for any function in real space. By expressing
a wavefunction from a system in the basis of the eigenstates of another system, one might
be able to learn about the nature of the wavefunction in question. For example, projecting
the wavefunction of a defect level in a periodic crystal onto the eigenfunctions of the
bulk crystal might yield information about how the defect level is constructed. For example,
the defect state might project primarily onto the conduction band, characterizing it
as a weakly perturbed host state. Such an understanding of defect levels could help select
corrections to perform on the defect system.\cite{lany}
For two structures (denoted R and S) with the same
lattice but different bases, the overlap operator of two Kohn-Sham single particle states
is described by the following terms:
\begin{equation}
\centering
\begin{split}
\braket{\psi_{Rn_1\mathbf{k}}|\psi_{Sn_2\mathbf{k}}} & =\widetilde{O}+O_M+O_R+O_S+O_N\\
O_M & =\sum_{a\in M_{RS}}\sum_{l,m}\sum_{\epsilon_1}\sum_{\epsilon_2}
\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|\widetilde{p}_{a\ind _1}}
(\braket{\phi_{a\ind _1}|{\phi_{a\ind _2}}}
-\braket{\widetilde{\phi}_{a\ind _1}|\widetilde{\phi}_{a\ind _2}})
\braket{\widetilde{p}_{a\ind _2}|\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_R & =\sum_{i\in N_R}\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|\widetilde{p}_i}
(\bra{\phi_i}-\bra{\widetilde{\phi}_i})
\ket{\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_S & =\sum_{j\in N_S}\bra{\widetilde{\psi}_{Rn_1\mathbf{k}}}(\ket{\phi_j}-\ket{\widetilde{\phi}_j})
\braket{\widetilde{p}_j|\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_N & =\sum_{i,j\in N_{RS}}\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|
\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})
(\ket{\phi_j}-\ket{\widetilde{\phi}_j})\braket{\widetilde{p}_j|\widetilde{\psi}_{Sn_2\mathbf{k}}}
\end{split}
\label{eq:finp}
\end{equation}
Here, $M_{RS}$ is the set of sites that are identical in structures R and S,
$N_R$ and $N_S$ are all other sites in these structures,
and $N_{RS}$ is teh set of all pairs of sites, one in structure R and one in structure
S, with overlapping augmentation spheres. Equation \ref{eq:finp} is derived in Appendix A.

@@

@@

\section{Projector function evaluation}

The projector function overlaps with the plane-wave states are calculated using the same
algorithm as VASP.\cite{vasp,vasp2,vasp3,vasp4}

\section{Projections for Overlapping Augmentation Spheres}

For two different structures, it is possible for atoms to have augmentation spheres
that overlap but are not directly on top of each other, or to have atoms of different elements
with the same position. In these cases, the overlap integrals of the partial waves
must be evaluated numerically on a per structure pair basis. The current method is a relatively
crude discrete integral in 3D space, which likely harms the precision of the overlap integral evaluation.
More precise and efficient methods are being considered.

@@

To test the program for its primary purpose of defect level analysis, the projection
of a defect level wavefunction onto the valence band of the bulk is defined:
\begin{equation}
v_d=\sum_{\mathbf{k}}\sum_{i\in \mathbb{V}}|\braket{\psi_{Bn_i\mathbf{k}}|\psi_{Dn_d\mathbf{k}}}|^2
\label{eq:vproj}
\end{equation}
where $i$ are the band indices of valence bands $\mathbb{V}$ in the bulk structure, the $B$
subscript denotes the bulk band structure, and the $D$ subscript denotes the point defect
band structure. This value, as well as the similarly defined $c_d$ for the conduction band,
were calculate for several charge states of the Silicon single vacancy, Phosphorus substitutional,
and Boron substitutional. The results are plotted in Figure \ref{fig:plots}.
As expected, the phosphorous and boron substitutional defects show litting mixing
of valence and conduction character in the defect levels because they are very shallow
defects, whereas the deeper vacancy character is shown by the mixing of conduction
and valence character in three nearly degenerate bands in the band gap.

\section{Conclusions}

PAWpySeed provides useful utilities for reading and analyzing PAW wavefunctions
from VASP, particularly for point defect levels. However, the current capabilities
of PAWpySeed are primarily qualitative because of mediocre numerical precision
and a lack of quantitative applications of the code. One potential quantitative
application to be developed in the future is the perturbation theory correction
proposedby Lany and Zunger for correcting defect level energies.\cite{lany}
The main barriers to such a correction are currently the lack of adequate
description of the deep level character of the defects and the difference
between the single particle states of DFT and higher level-of-theory
single-particle states. Work to resolve such issues is currently in development.
Namely, code is being written to construct atomic vacuum states onto which
PAW wavefunctions can be projected. In addition, the SCDM-k decomposition method
will written into PAWpySeed to construct localized orbitals for energy
corrections.\cite{scdm,scdmk}

@@

The starting point for this derivation is the transformation operator that
is performed on the pseudowavefunction to obtain the all electron wavefunction
\begin{equation}
T=1+\sum_a\sum_l\sum_m\sum_{\epsilon}(\ket{\phi_{a\ind}}-\ket{\widetilde{\phi}_{a\ind}})
\bra{\widetilde{p}_{a\ind}}=1+\sum_i(\ket{\phi_i}-\ket{\widetilde{\phi}_i})
\bra{\widetilde{p}_i}
\label{eq:T}
\end{equation}
where $\phi_{a\ind}$ are all electron (AE) partial waves, $\widetilde{\phi}_{a\ind}$
are pseudo (PS) partial waves, $\widetilde{p}_{a\ind}$ are projector functions,
$a$ are the site indices of each atom in the structure, and $l$, $m$, and $\epsilon$
specify a spherical harmonic and index which uniquely specify partial wave at a given
index. A summation over $i$ (or $j$, as below) represents a summation over $a$, $l$, $m$,
and $\epsilon$. For further details on the PAW method, including the physical significance
and construction of the partial waves and projector functions, see Blochl's original paper
and Kresse and Joubert's paper relating ultrasoft pseudopotentials and PAW.\cite{blochl,kresse_joubert}
The next step is to define a pseudo operator $\widetilde{A}$ for each operator $A$
such that $\bra{\psi}A\ket{\psi}=\bra{\widetilde{\psi}}\widetilde{A}\ket{\widetilde{\psi}}$.
Because $\ket{\psi}=T\ket{\widetilde{\psi}}$, one can write
\begin{equation}
\widetilde{A}=T^{\dagger}AT
\label{eq:op}
\end{equation}
One can then plug Equation \ref{eq:T} into Equation \ref{eq:op} to find
\begin{equation}
\begin{split}
\widetilde{A} = & [1+\sum_i\ket{\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})]
A[1+\sum_j(\ket{\phi_j}-\ket{\widetilde{\phi}_j})
\bra{\widetilde{p}_j}]\\
\widetilde{A} = & A+\sum_i\ket{\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})A
+\sum_j A(\ket{\phi_j}-\ket{\widetilde{\phi}_j})\bra{\widetilde{p}_j}\\
& +\sum_i\sum_j\ket{\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})A
(\ket{\phi_j}-\ket{\widetilde{\phi}_j})\bra{\widetilde{p}_j}
\end{split}
\label{eq:res}
\end{equation}

@@

one can rearrange Equation \ref{eq:res} in the manner of Blochl\cite{blochl}:
\begin{equation}
\begin{split}
\widetilde{A}= & A+\sum_i\sum_j[\ket{\widetilde{p}_i}(\bra{\phi_i}A\ket{\phi_j}
-\bra{\widetilde{\phi_i}}A\ket{\widetilde{\phi_j}})\bra{\widetilde{p}_j}] \\
& +\sum_i[(1-\sum_j\ket{\widetilde{p}_j}\bra{\widetilde{\phi}_j})A(\ket{\phi_i}-\ket{\widetilde{\phi_i}})
\bra{\widetilde{p}_i}+\ket{\widetilde{p}_i}
(\bra{\phi_i}-\bra{\widetilde{\phi}_i})A(1-\sum_j\ket{\widetilde{\phi}_j}\bra{\widetilde{p}_j})]
\end{split}
\label{eq:nonloc}
\end{equation}
When the operator $A$ is local, then $\sum_j\ket{\widetilde{\phi}_j}\bra{\widetilde{p}_j}=1$,
so the entire second line of Equation \ref{eq:nonloc} vanishes, giving Blochl's formulation:\cite{blochl}
\begin{equation}
\widetilde{A}=A+\sum_i\sum_j\ket{\widetilde{p}_i}(\bra{\phi_i}A\ket{\phi_j}
-\bra{\widetilde{\phi_i}}A\ket{\widetilde{\phi_j}})\bra{\widetilde{p}_j}
\label{eq:loc}
\end{equation}

However, while the overlap operator is local, the simplification in Equation \ref{eq:loc} assumes that
the summations over $i$ and $j$ go over the same species at the same locations in the same lattice;
only the third condition is always satisfied for systems of interest for projections between different
bases. Equation \ref{eq:res} is a more convenient representation than Equation \ref{eq:nonloc} for this
purpose because it is easily seen that all terms between projectors whose augmentation regions do not
overlap vanish. In addition, the POTCAR file in VASP only stores the PS and AE partial waves out to the
edge of the augmentation region, where they are equal but not nonzero, so it is useful to organize the terms
so that they are guaranteed to vanish outside the augmentation regions. Replacing $A$ with unity
allows the overlap between two bands
in different structures $R$ and $S$ with the same lattice to be specified (note that the
summation over $i$ is for structure $R$ and the summation over $j$ is for structure $S$):
\begin{equation}
\begin{split}
\braket{\psi_{Rn_1\mathbf{k}}|\psi_{Sn_2\mathbf{k}}} & =\widetilde{O}+O_1+O_2+O_3\\
\widetilde{O} & =\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_1 & =\sum_i\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})
\ket{\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_2 & =\sum_j\bra{\widetilde{\psi}_{Rn_1\mathbf{k}}}(\ket{\phi_j}-\ket{\widetilde{\phi}_j})
\braket{\widetilde{p}_j|\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_3 & =\sum_i\sum_j\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|
\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})
(\ket{\phi_j}-\ket{\widetilde{\phi}_j})\braket{\widetilde{p}_j|\widetilde{\psi}_{Sn_2\mathbf{k}}}
\end{split}
\label{eq:ov1}
\end{equation}

>>>>

\documentclass[12pt]{article}
\usepackage{authblk}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{float}
\usepackage{braket}
\usepackage[margin=1.0in]{geometry}
\usepackage[style=chem-acs,backend=bibtex,articletitle=false]{biblatex}
\graphicspath{ {../../Si_stuff/} }
\bibliography{mybib}

@@

\title{PAWpySeed: A parallelized Python/C package for analyzing
DFT wavefunctions in the PAW formalism}
\author[1]{Kyle Bystrom (kylebystrom@berkeley.edu)}
\affil[1]{UC Berkeley Department of Materials Science and Engineering}
\date{\today}
\maketitle
\abstract{\emph{Note: My current plan is to build this work into a senior thesis,
but currently it is quite messy and incomplete because I still have a lot of work
to do before I will feel comfortable presenting this as a complete work. I would appreciate
any feedback on what I have done so far.}

A parallelized Python and C package has been developed
to perform analysis of density functional theory (DFT) wavefunctions in the
projector augmented wave (PAW) formalism. The primary function of this code
is to calculate the overlap operator of two single-particle Kohn-Sham states
from different periodic structures with the same lattice, which is made nontrivial
by the dependency of PAW augmentation regions on the basis of the structure.
The code is potentially useful for the analysis of point defect levels in
solids because it can determine the composition of defect levels in the basis
of the single-particle states of the bulk system.}

\section{Overview}

In the pure physical picture of a quantum mechanical system, the eigenfunctions of the time
independent Hamiltonian form a complete basis for any function in real space. By expressing
a wavefunction from a system in the basis of the eigenstates of another system, one might
be able to learn about the nature of the wavefunction in question. For example, projecting
the wavefunction of a defect level in a periodic crystal onto the eigenfunctions of the
bulk crystal might yield information about how the defect level is constructed. For example,
the defect state might project primarily onto the conduction band, characterizing it
as a weakly perturbed host state. Such an understanding of defect levels could help select
corrections to perform on the defect system.\cite{lany}

The program developed here is designed to evaluate the overlap operators necessary for the
above projections in the projector augmented wave formalism (PAW) developed by Blochl.\cite{blochl}
pawpyseed interfaces with the Vienna \emph{ab initio} Simulaton
Package (VASP)\cite{vasp,vasp1,vasp2,vasp3,vasp4}, but will interface with other plane wave
codes in the future.

The main purpose of this code is to analyze defect wavefunctions, as explained above;
however, the formalism adn interface developed is relatively general in purpose, and
may suit other uses. The code can perform convenient operations not included in VASP,
such as writing all electron wavefunctions with phases to volumetric data files, which may
be useful for visualization. In addition, one future application plan is to use a given
set of norm-conserving pseudopotentials to convert PAW wavefunctions to NC wavefunctions,
which could be useful for researchers working with GW codes, which usually require
NC wavefunctions. Lastly, operators besides overlap operators can hypothetically be evaluated
in this formalism, which might be useful for optical applications, since materials can
change structure when absorbing a photon.

@@

The projector augmented wave (PAW) method defines a transformation between the all electron
(AE) single particle states of the Kohn Sham wavefunction and a pseudo (PS) wavefunction
that is smooth throughout the lattice and therefore can be expressed in a plane wave basis set.\cite{blochl}
This means that the AE wavefunction is a sum of plane waves and site-localized atomic states,
which makes calculating operators difficult if the ion positions change.

@@

\subsubsection{Projection onto Basis from Different Structure}
For two structures (denoted R and S) with the same
lattice but different bases, the overlap operator of two Kohn-Sham single particle states
is described by the following terms:
\begin{equation}
\centering
\begin{split}
\braket{\psi_{Rn_1\mathbf{k}}|\psi_{Sn_2\mathbf{k}}} & =\widetilde{O}+O_M+O_R+O_S+O_N\\
O_M & =\sum_{a\in M_{RS}}\sum_{l,m}\sum_{\epsilon_1}\sum_{\epsilon_2}
\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|\widetilde{p}_{a\ind _1}}
(\braket{\phi_{a\ind _1}|{\phi_{a\ind _2}}}
-\braket{\widetilde{\phi}_{a\ind _1}|\widetilde{\phi}_{a\ind _2}})
\braket{\widetilde{p}_{a\ind _2}|\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_R & =\sum_{i\in N_R}\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|\widetilde{p}_i}
(\bra{\phi_i}-\bra{\widetilde{\phi}_i})
\ket{\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_S & =\sum_{j\in N_S}\bra{\widetilde{\psi}_{Rn_1\mathbf{k}}}(\ket{\phi_j}-\ket{\widetilde{\phi}_j})
\braket{\widetilde{p}_j|\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_N & =\sum_{i,j\in N_{RS}}\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|
\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})
(\ket{\phi_j}-\ket{\widetilde{\phi}_j})\braket{\widetilde{p}_j|\widetilde{\psi}_{Sn_2\mathbf{k}}}
\end{split}
\label{eq:finp}
\end{equation}
Here, $M_{RS}$ is the set of sites that are identical in structures R and S,
$N_R$ and $N_S$ are all other sites in these structures,
and $N_{RS}$ is teh set of all pairs of sites, one in structure R and one in structure
S, with overlapping augmentation spheres. Equation \ref{eq:finp} is derived in Appendix A.

@@

\subsection{Projector function evaluation}

The projector function overlaps with the plane-wave states are calculated using the same
algorithm as VASP.\cite{vasp}

@@

The current implementation uses the "high-k" algorithm developed by Talman. At low
values of k, this algorithm becomes imprecise, though the precision
of the current implementation has been tested and found to be about
$10^{-3}$ for a bulk Ga test system. Future versions of the code
will select the most precise of the "low-k" and "high-k" algorithms for each point,
and the density of the momentum space grid will be increased.

The main problem with this algorithm is that it is the slowest component of the code
and consumes the majority of the computational time even when only a few sites out
of a structure with hundreds of sites require the calculation. A potential fix is,
rather than integrating in reciprocal space, which becomes prohibitively slow for large
lattices, to remove the high-frequency components of the partial waves via Fourier transform.
This is counterintuitive because the purpose of the partial waves is to preserve the high-frequency
components of the atomic wavefunctions, but when projecting onto the smooth pseudowavefunction,
the high-frequency components have no impact because they are not present in the pseudowavefunction.
Removing the high frequency components, therefore, allows the projection onto the partial waves
to be done in real space without aliasing.

\subsection{Projections for Overlapping Augmentation Spheres}

For two different structures, it is possible for atoms to have augmentation spheres
that overlap but are not directly on top of each other, or to have atoms of different elements
with the same position. In these cases, the overlap integrals of the partial waves
must be evaluated numerically on a per structure pair basis. The current method is a relatively
crude discrete integral in 3D space, which is the limiting factor in the precision of the code.
The current precision of this two-center overlap integral is about $10^{-2}$, which is adequate
for qualitative evaluations of band composition but not for quantitative charge corrections.

An implementation is in development in which a kernel function developed by Hierse and
Oppeneer is used to reduce the 3-dimensional discrete integrals
to the integral of an analytical term over the two radial distribution functions.\cite{hierse_kernel} 
This method has the potential to vastly improve both the performance and precision of
this step of the code, but the relevant kernel function for overlap integrals
scales inversely with the internuclear separation, so detailed tests are required
to design the implementation in a numerically precise fashion. Most likely, it will be
best to divide the integrals into three cases: (1) internuclear separation is very small
and can be assumed to vanish, (2) internuclear separation is small, but not small enough
to be ignored numerically, but the kernel function at a given $r_1$ is very narrow
in $r_2$, (3) the internuclear separation is large and the kernel function smooth everywhere
relative to the partial wave grid density.

The Gaunt coefficients are required for the evaluation of the kernel function, and these
are calculated and stored efficiently with sympy and the selection rules for \cite{gaunt_selection}.

@@

To test the program for its primary purpose of defect level analysis, the projection
of a defect level wavefunction onto the valence band of the bulk is defined:
\begin{equation}
v_d=\sum_{\mathbf{k}}\sum_{i\in \mathbb{V}}|\braket{\psi_{Bn_i\mathbf{k}}|\psi_{Dn_d\mathbf{k}}}|^2
\label{eq:vproj}
\end{equation}
where $i$ are the band indices of valence bands $\mathbb{V}$ in the bulk structure, the $B$
subscript denotes the bulk band structure, and the $D$ subscript denotes the point defect
band structure. This value, as well as the similarly defined $c_d$ for the conduction band,
were calculated for several charge states of the silicon single vacancy, phosphorus substitutional,
and boron substitutional. A 128-atom supercell was used (127 for the vacancy), with
the PBE GGA functional in VASP.\cite{gga,ggaerr} A Monkhorst-Pack 2x2x2 k-point mesh was used.

\begin{figure}
\includegraphics[width=0.32\textwidth]{Bcharge_0.png}
\includegraphics[width=0.32\textwidth]{Bcharge_-1.png}
\includegraphics[width=0.32\textwidth]{Pcharge_0.png}
\includegraphics[width=0.32\textwidth]{Pcharge_1.png}
\includegraphics[width=0.32\textwidth]{charge_-2.png}
\includegraphics[width=0.32\textwidth]{charge_-1.png}
\includegraphics[width=0.32\textwidth]{charge_0.png}
\includegraphics[width=0.32\textwidth]{charge_1.png}
\includegraphics[width=0.32\textwidth]{charge_2.png}
\caption{Valence and conduction band character of several charge states
of three Si point defects, determined by evaluating the overlap
operator between all electron wavefunctions. The charge states prefixed by B are
boron substitutionals, those prefixed by P are phosphorous substitutionals,
and those without a prefix are Si single vacancies.}
\label{fig:ae_proj}
\end{figure}

\begin{figure}
\includegraphics[width=0.32\textwidth]{pseudo/Bcharge_0.png}
\includegraphics[width=0.32\textwidth]{pseudo/Bcharge_-1.png}
\includegraphics[width=0.32\textwidth]{pseudo/Pcharge_0.png}
\includegraphics[width=0.32\textwidth]{pseudo/Pcharge_1.png}
\includegraphics[width=0.32\textwidth]{pseudo/charge_-2.png}
\includegraphics[width=0.32\textwidth]{pseudo/charge_-1.png}
\includegraphics[width=0.32\textwidth]{pseudo/charge_0.png}
\includegraphics[width=0.32\textwidth]{pseudo/charge_1.png}
\includegraphics[width=0.32\textwidth]{pseudo/charge_2.png}
\caption{Valence and conduction band character of several charge states
of three Si point defects, determined by only evaluating
the overlap operator between pseudowavefunctions and then forcibly
normalizing the result to 1. The charge states prefixed by B are
boron substitutionals, those prefixed by P ar phosphorous substitutionals,
and those without a prefix are Si single vacancies.}
\label{fig:ps_proj}
\end{figure}

Figure \ref{fig:ae_proj} shows the valence and conduction band characters of
each defect and charge state. The shallow character of the phosphorous and boron
substitutionals result in each band being strongly associated with either the bulk
or conduction bands. In the Si vacancy, three bands have some mixing of conduction
and valence character, which is sensible because the vacancy is a deeper defect.

The same defects are analyzed in Figure \ref{fig:ps_proj}, except that the overlap
operators were evaluated using pseudowavefunctions, which is much faster but
less physically meaningful because the pseudowavefunctions are unnormalized
and not orthogonal (the
diagram forcibly normalizes valence and conduction character) and
different in shape than the true AE single-particle states. However, as shown
for these simple Si systems, the same qualitative valence-conduction character is
observed, indicating that such qualitative methods might be useful for more quickly
analyzing bands.

The sulfur substitutional in silicon was analyzed in a 216-atom supercell,
but further work is required to interpret the results meaningfully, so they
will be released in future work.

\section{Conclusions and Future Work}

Most of the work performed thus far has centered on the writing and testing of
PAWpySeed for basic functionality on various systems. As elaborated below,
a major goal for future work is to add new functionality to PAWpySeed and increase
its precision. However, the most prescient matter is taking the results
accumulated on test defect systems and comparing the output of PAWpySeed to the
DFT formation energies and eigenvalues of the systems. The results above
verify the functionality of the code but not its usefulness, which will only
be discernible in comparision with the quantitative output of the DFT calculations.
This, along with increasing the efficiency and precision of the code, will be
the main goal of work over the next several months.

PAWpySeed provides useful utilities for reading and analyzing PAW wavefunctions
from VASP, particularly for point defect levels. However, the current capabilities
of PAWpySeed are primarily qualitative because of mediocre numerical precision
and a lack of quantitative applications of the code. One potential quantitative
application to be developed in the future is the perturbation theory correction
proposed by Lany and Zunger for correcting defect level energies.\cite{lany}
The main barriers to such a correction are currently the lack of adequate
description of the deep level character of the defects and the difference
between the single particle states of DFT and higher level-of-theory
single-particle states. Work to resolve such issues is currently in development.
Namely, code is being written to construct atomic vacuum states onto which
PAW wavefunctions can be projected. In addition, the SCDM-k decomposition method
will be written into PAWpySeed to construct localized orbitals for energy
corrections.\cite{scdm,scdmk}

\section{Acknowledgments}

I would like to thank Danny Broberg and Mark Asta for excellent advising
and encouragement throughout my work with the Asta Research Group at UC Berkeley,
as well as all the members of the Asta Research Group and Hacking Materials
Research Group at Lawrence Berkeley National Laboratory for various
valuable research and learning opportunities. My summer 2017 undergraduate
research, mainly related to development of an open-source data science code
called matminer, was supported
by a summer research stipend from the UC Berkeley College of Chemistry.

@@

The starting point for this derivation is the transformation operator that
is performed on the pseudowavefunction to obtain the all electron wavefunction
\begin{equation}
T=1+\sum_a\sum_l\sum_m\sum_{\epsilon}(\ket{\phi_{a\ind}}-\ket{\widetilde{\phi}_{a\ind}})
\bra{\widetilde{p}_{a\ind}}=1+\sum_i(\ket{\phi_i}-\ket{\widetilde{\phi}_i})
\bra{\widetilde{p}_i}
\label{eq:T}
\end{equation}
where $\phi_{a\ind}$ are all electron (AE) partial waves, $\widetilde{\phi}_{a\ind}$
are pseudo (PS) partial waves, $\widetilde{p}_{a\ind}$ are projector functions,
$a$ are the site indices of each atom in the structure, and $l$, $m$, and $\epsilon$
specify a spherical harmonic and index which uniquely specify partial wave at a given
index. A summation over $i$ (or $j$, as below) represents a summation over $a$, $l$, $m$,
and $\epsilon$. For further details on the PAW method, including the physical significance
and construction of the partial waves and projector functions, see Blochl's original paper
and Kresse and Joubert's paper relating ultrasoft pseudopotentials and PAW.\cite{blochl,vasp4}
The next step is to define a pseudo operator $\widetilde{A}$ for each operator $A$
such that $\bra{\psi}A\ket{\psi}=\bra{\widetilde{\psi}}\widetilde{A}\ket{\widetilde{\psi}}$.
Because $\ket{\psi}=T\ket{\widetilde{\psi}}$, one can write
\begin{equation}
\widetilde{A}=T^{\dagger}AT
\label{eq:op}
\end{equation}
One can then plug Equation \ref{eq:T} into Equation \ref{eq:op} to find
\begin{equation}
\begin{split}
\widetilde{A} = & [1+\sum_i\ket{\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})]
A[1+\sum_j(\ket{\phi_j}-\ket{\widetilde{\phi}_j})
\bra{\widetilde{p}_j}]\\
\widetilde{A} = & A+\sum_i\ket{\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})A
+\sum_j A(\ket{\phi_j}-\ket{\widetilde{\phi}_j})\bra{\widetilde{p}_j}\\
& +\sum_i\sum_j\ket{\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})A
(\ket{\phi_j}-\ket{\widetilde{\phi}_j})\bra{\widetilde{p}_j}
\end{split}
\label{eq:res}
\end{equation}

@@

one can rearrange Equation \ref{eq:res} in the manner of Blochl\cite{blochl}:
\begin{equation}
\begin{split}
\widetilde{A}= & A+\sum_i\sum_j[\ket{\widetilde{p}_i}(\bra{\phi_i}A\ket{\phi_j}
-\bra{\widetilde{\phi_i}}A\ket{\widetilde{\phi_j}})\bra{\widetilde{p}_j}] \\
& +\sum_i[(1-\sum_j\ket{\widetilde{p}_j}\bra{\widetilde{\phi}_j})A(\ket{\phi_i}-\ket{\widetilde{\phi_i}})
\bra{\widetilde{p}_i}+\ket{\widetilde{p}_i}
(\bra{\phi_i}-\bra{\widetilde{\phi}_i})A(1-\sum_j\ket{\widetilde{\phi}_j}\bra{\widetilde{p}_j})]
\end{split}
\label{eq:nonloc}
\end{equation}
When the operator $A$ is local, then $\sum_j\ket{\widetilde{\phi}_j}\bra{\widetilde{p}_j}=1$,
so the entire second line of Equation \ref{eq:nonloc} vanishes, giving Blochl's formulation:\cite{blochl}
\begin{equation}
\widetilde{A}=A+\sum_i\sum_j\ket{\widetilde{p}_i}(\bra{\phi_i}A\ket{\phi_j}
-\bra{\widetilde{\phi_i}}A\ket{\widetilde{\phi_j}})\bra{\widetilde{p}_j}
\label{eq:loc1}
\end{equation}

However, while the overlap operator is local, the simplification in Equation \ref{eq:loc1} assumes that
the summations over $i$ and $j$ go over the same species at the same locations in the same lattice;
only the third condition is always satisfied for systems of interest for projections between different
bases. Equation \ref{eq:res} is a more convenient representation than Equation \ref{eq:nonloc} for this
purpose because it is easily seen that all terms between projectors whose augmentation regions do not
overlap vanish. In addition, the POTCAR file in VASP only stores the PS and AE partial waves out to the
edge of the augmentation region, where they are equal but not nonzero, so it is useful to organize the terms
so that they are guaranteed to vanish outside the augmentation regions. Replacing $A$ with unity
allows the overlap between two bands
in different structures $R$ and $S$ with the same lattice to be specified (note that the
summation over $i$ is for structure $R$ and the summation over $j$ is for structure $S$):
\begin{equation}
\begin{split}
\braket{\psi_{Rn_1\mathbf{k}}|\psi_{Sn_2\mathbf{k}}} & =\widetilde{O}+O_1+O_2+O_3\\
\widetilde{O} & =\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_1 & =\sum_i\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})
\ket{\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_2 & =\sum_j\bra{\widetilde{\psi}_{Rn_1\mathbf{k}}}(\ket{\phi_j}-\ket{\widetilde{\phi}_j})
\braket{\widetilde{p}_j|\widetilde{\psi}_{Sn_2\mathbf{k}}}\\
O_3 & =\sum_i\sum_j\braket{\widetilde{\psi}_{Rn_1\mathbf{k}}|
\widetilde{p}_i}(\bra{\phi_i}-\bra{\widetilde{\phi}_i})
(\ket{\phi_j}-\ket{\widetilde{\phi}_j})\braket{\widetilde{p}_j|\widetilde{\psi}_{Sn_2\mathbf{k}}}
\end{split}
\label{eq:ov1a}
\end{equation}
